### 2020-05-25 manuscript/manuscript.rmd
modify main figures and legends

To advance this problem it is therefore critical to synthesize published results and explore variation in a standardized framework. Here, we synthesize data and results from previous phylogenomic analyses that tested Ctenophora-sister and Porifera-sister from 14 studies (Table 1). These matrices include all relevant studies published before 2020 for which we could obtain data matrices with gene partition annotations.

@@

The published analyses considere here differ in a variety of ways, including the model of molecular evolution, gene selection, ingroup species selection, and outgroup species selection.

@@

Sampling of ingroup taxa (animals) has always been heavilly biased toward Bilateria, but sampling of non-bilaterian groups, including sponges and ctenophores, has improved over time as studies add additional taxa. As sequencing technologies have developed, there has also been some differences across studies in which members of particular groups are included as older shallower datasets are replaced with new deeper datasets from closely related species. Ingroup taxon sampling across clades varies widely between studies (Extended Data Fig. 1). The first studies that use new sequencing technologies tend to have a greater fraction of species belonging to Bilateria, since it includes many of the best-studied animal groups and all caononical model species where these technologies are first deployed. More than 50% taxa sampled were Bilateria in the earliest transcriptome (Dunn2008 [@Dunn:2008ky] and Hejnol2009 [@hejnol2009assessing]) and genome genome (Borrowiec2015 [@borowiec2015extracting]) studies considered here. Importantly, several earlier studies before 2015 only include less than five ctenophore species (Extended Data Fig. 1). Within each clade, there is often considerable variation in taxon sampling and therefore little species overlap across studies (Extended Data Fig. XXXX).

@@

no_subsets = unlist(lapply(sequence_matrices, function(x) !grepl("only|_D", x@matrix_name)))

@@

taxon_rectangles$matrix[taxon_rectangles$matrix == "Simion2017_subsampled"] <- "Simion2017"
taxon_rectangles$matrix[taxon_rectangles$matrix == "Hejnol2009_subsampled"] <- "Hejnol2009"

clade_colors = c("#e31a1c","#a6cee3","#1f78b4","#b2df8a","#33a02c","#fb9a99","#a6cee3","#fdbf6f","#cab2d6", "#ff7f00")
names( clade_colors ) = clades

@@

@@

**Fig. XXtaxon_rectangles.** Matrix taxa and gene sampling. Each of the primary matrices considered here, color coded by taxon sampling. Horizontal size is proportional to the number of genes sampled, vertical size to the number of taxa sampled.

@@

```{r figure_support_analyses_published}
aanalyses_published=read_tsv("../data_processed/tables/previously_published_analyses.tsv")
analyses_published_filter = dplyr::filter(analyses_published, !grepl('Recoding', model_combined))
x1 = factor(analyses_published_filter$model_combined, levels=c("WAG", "LG", "GTR", "data partitioning", "Recoding + GTR", "Recoding + GTR + CAT", "Poisson + CAT", "GTR + CAT"))
p1 = analyses_published_filter %>%
	ggplot( aes(x=x1, y=clade, col=final, shape=inference) ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))+
   xlab("Model complexity") + ylab("Clade")

@@

x2 = factor(analyses_new$model, levels=c("WAG", "GTR", "Poisson+C60", "WAG+C60", "LG+C60", "Poisson+CAT"))
p2 = analyses_new %>%
	ggplot( aes(x=x2, y=clade, col=result, shape=inference) ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))+
       xlab("Model complexity") + ylab("Clade")

@@

```
**Fig. XXfigure_support_analyses.**  A summary of phylogenomic results from the previous studies and reanalyses conducted in this study. (A) A total of `r nrow(analyses_published_filter)` analyses were transcribed from the literature (Table XX). (B) A total of `r nrow(analyses_new)` new phylogenomic analyses were conducted in this study (Supplementary Table XX).

@@

The distribution of phylogenetic signal has never been quantified in site-heterogeneous models and compared with site-homogeneous models. To further explore the effect of model selection and outgroup choices in animal origin, we next quantify the support of phylogenetic signal over two alternative hypotheses (T1: Ctenophora-sister (Fig. 1A); T2: Porifera-sister (Fig. 1B)) to three representative data matrices with different outgroup choice and models in both ML and BI framework (Fig. 5). By calculating gene-wise log-likelihood scores between T1 and T2 for every gene ($\Delta$GLS) or site ($\Delta$SLS) in each matrix, we found that Ctenophora-sister had the higher proportions of supporting genes in every analysis when using site-homogeneous models (LG or WAG in both IQtree or PhyloBayes). Moreover, the outgroup choice has little impact on the distribution of the support of phylogenetic signal in sifte-homogeneous models. This finding is largely consistent with the results from past study that majority of phylogenetic signal is strongly favored in Ctenophora-sister hypothesis with site-homogeneous models in other data matrices [@shen2017contentious].

@@

We next explore how the increase of number of categories may affect the phylogenetic inference related to the animal phylogeny with a panel of different substitutional categories in the representative data matrices (Fig. XX). Surprisingly, we found Poisson+nCAT60 showed exact same topology as Poisson+C60 or WAG+C60 with strong support of Ctenophora-sister, suggesting that the different phylogenetic analysis software (IQtree and PhyloBayes) and phylogenetic inference (ML and BI) has limited impact on phylogenetic inference to these matrices with similar evolutioanry models. Importantly, our results gave decreased support for Ctenophora-sister and increased support of Porifera-sister with the increase of number of substitutional categories (Fig. XX). More importantly, we are able to identify the transition of topology between the two alternative hypotheses in these three matrices. For example, the major transition occurred at 60 -- 120 categories, and the support for Ctenophora-sister disappears after 120 categories in the Whelan2017_strict matrix. Interestingly, the results of our sensitive analyses further corroborated by the recent simulation study that the tendency of overestimate the true number of categories of Poisson-CAT or GTR-CAT model can potentially lead to a less accurate tree than the site-homogeneous models [@whelan2016let].

>>>>

To advance this problem it is therefore critical to synthesize published results and explore variation in a standardized framework. Here, we synthesize data and results from previous phylogenomic analyses that tested Ctenophora-sister and Porifera-sister from 13 studies (Table 1). These matrices include all relevant studies published before 2020 for which we could obtain data matrices with gene partition annotations.

@@

The published analyses considered here differ in a variety of ways, including the model of molecular evolution, gene selection, ingroup species selection, and outgroup species selection.

@@

@@

no_subsets = unlist(lapply(sequence_matrices, function(x) !grepl("only", x@matrix_name)))

@@

taxon_rectangles$matrix[taxon_rectangles$matrix == "Simion2017_subsampled"] <- "Simion2017"
taxon_rectangles$matrix[taxon_rectangles$matrix == "Hejnol2009_subsampled"] <- "Hejnol2009"
clade_colors = c("#e31a1c","#fb9a99","#fb9a99","#1e90ff","#E69F00","#009E73","#1f78b4","#cab2d6","#c71585")
names( clade_colors ) = clades

taxon_rectangles= taxon_rectangles %>% filter(matrix!="Nosenko2013_nonribo_9187") %>% filter(matrix!="Whelan2017_strict") %>% filter(matrix!="Whelan2015_D1") %>% filter(matrix!="Whelan2015_D10") %>% filter(matrix!="Nosenko2013_ribo_11057")%>% filter(matrix!="Borowiec2015_Best108")

@@

	#matrix = factor(taxon_rectangles$matrix, levels=c("Dunn2008", "Hejnol2009", "Philippe2009", "Nosenko2013_ribo_14615", "Ryan2013_est", "Moroz2014_3d", "Whelan2015_D20", "Borowiec2015_Total1080" , "Chang2015", "Whelan2017_full", "Simion2017"))

@@

**Fig. XXtaxon_rectangles.** An overview of previous studies. (A) A time-series plot to show the main conclusions of previous 13 phylogenomic studies that tested Ctenophora-sister (designated by the orange node) and Porifera-sister (designated by the green node). (B) Matrix taxa and gene sampling for each study. Each of the primary matrices considered here, color coded by taxon sampling. Horizontal size is proportional to the number of genes sampled, vertical size to the number of taxa sampled.

Sampling of ingroup taxa (animals) has always been heavilly biased toward Bilateria, but sampling of non-bilaterian groups, including sponges and ctenophores, has improved over time as studies add additional taxa. As sequencing technologies have developed, there has also been some differences across studies in which members of particular groups are included as older shallower datasets are replaced with new deeper datasets from closely related species. Ingroup taxon sampling across clades varies widely between studies (figure_taxon_rectangles). The first studies that use new sequencing technologies tend to have a greater fraction of species belonging to Bilateria, since it includes many of the best-studied animal groups and all caononical model species where these technologies are first deployed. More than 50% taxa sampled were Bilateria in the earliest transcriptome (Dunn2008 [@Dunn:2008ky] and Hejnol2009 [@hejnol2009assessing]) and genome genome (Borrowiec2015 [@borowiec2015extracting]) studies considered here. Importantly, several earlier studies before 2015 only include less than five ctenophore species (Extended Data Fig. 1). Within each clade, there is often considerable variation in taxon sampling and therefore little species overlap across studies (Extended Data Fig. XXXX).

@@

```{r figure_support_analyses_published}
analyses_published=read_tsv("../data_processed/tables/previously_published_analyses.tsv")
analyses_published_filter = dplyr::filter(analyses_published, !grepl('Recoding', model_combined))
x1 = factor(analyses_published_filter$model_combined, levels=c("WAG", "LG", "GTR", "data partitioning", "Recoding + GTR", "Recoding + GTR + CAT", "Poisson + CAT", "GTR + CAT"))
p1 = analyses_published_filter %>%
	ggplot( aes(x=x1, y=clade, col=final, shape=inference) ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))+
   xlab("Model complexity") + ylab("Clade")

@@

x2 = factor(analyses_new$model, levels=c("WAG", "GTR", "Poisson+C60", "WAG+C60", "LG+C60", "Poisson+CAT", "GTR+CAT"))
p2 = analyses_new %>%
	ggplot( aes(x=x2, y=clade, col=result, shape=inference) ) +
		geom_jitter( width = 0.20, height = 0.20, alpha=0.7, size=2 ) +
		scale_colour_manual(values = result_colors) +
		theme_classic() + theme(axis.text.x = element_text(angle=45, hjust=1))+
       xlab("Model complexity") + ylab("Clade")

@@

```
**Fig. XXfigure_support_analyses.**  A summary of phylogenomic results from the previous studies and reanalysis conducted in this study. Each shape represents an phylogenetic analysis that has been analyzed by ML (designated by the circle) or BI inference (designated by the triangle). Analyses with a support of bootstrap values less than 90% or a posterior probability less than 0.95 are considered as unresolved. (A) A total of `r nrow(analyses_published_filter)` analyses were transcribed from the literature (related to Supplementary Table XX). (B) A total of `r nrow(analyses_new)` new phylogenomic analyses were conducted in this study (related to Supplementary Table XX).

@@

The distribution of phylogenetic signal has never been quantified in site-heterogeneous models and compared with site-homogeneous models. To further explore the effect of model selection and outgroup choices in animal origin, we next quantify the support of phylogenetic signal over two alternative hypotheses (T1: Ctenophora-sister (Fig. 1A); T2: Porifera-sister (Fig. 1B)) to three representative data matrices with different outgroup choice and models in both ML and BI framework (Fig. 5). By calculating gene-wise log-likelihood scores between T1 and T2 for every gene ($\Delta$GLS) in each matrix, we found that Ctenophora-sister had the higher proportions of supporting genes in every analysis when using site-homogeneous models (LG or WAG in both IQtree or PhyloBayes). Moreover, the outgroup choice has little impact on the distribution of the support of phylogenetic signal in sifte-homogeneous models. This finding is largely consistent with the results from past study that majority of phylogenetic signal is strongly favored in Ctenophora-sister hypothesis with site-homogeneous models in other data matrices [@shen2017contentious].

@@

We next explore how the increase of number of categories may affect the phylogenetic inference related to the animal phylogeny with a panel of different substitutional categories in the representative data matrices (Fig. XX). Surprisingly, we found Poisson+nCAT60 showed exact same topology as Poisson+C60 or WAG+C60 with strong support of Ctenophora-sister, suggesting that the different phylogenetic analysis software (IQtree and PhyloBayes) and phylogenetic inference (ML and BI) has limited impact on phylogenetic inference to these matrices with similar evolutionary models. Importantly, our results gave decreased support for Ctenophora-sister and increased support of Porifera-sister with the increase of number of substitutional categories (Fig. XX). More importantly, we are able to identify the transition of topology between the two alternative hypotheses in these three matrices. For example, the major transition occurred at 60 -- 120 categories, and the support for Ctenophora-sister disappears after 120 categories in the Whelan2017_strict matrix. Interestingly, the results of our sensitive analyses further corroborated by the recent simulation study that the tendency of overestimate the true number of categories of Poisson-CAT or GTR-CAT model can potentially lead to a less accurate tree than the site-homogeneous models [@whelan2016let].
